Evolution of CSMA Protocols of the IEEE 802.11 Standard

### 2013-01-22 paper.tex
working on related work

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

@@

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Evolution of CSMA Protocols fro the IEEE 802.11 Standard}

@@

A candidate to replace CSMA/CA should 
\begin{itemize}
  \item Provide performance advantages, either in the form of throughput or short term fairness.
  \item Be backward compatible with current implementation.
  \item Be simple a simple evolution implementation to ease the transition and reduce time to market (Optional but desirable).
\end{itemize}
\section{Related Work}

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

>>>>

% *** CITATION PACKAGES ***
%
\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

@@

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Evolution of CSMA Protocols of the IEEE 802.11 Standard}

@@

A candidate to replace CSMA/CA should 
\begin{itemize}
  \item Provide performance advantages, either in the form of throughput or short term fairness.
  \item Be backward compatible with current implementation.
  \item Be simple a simple evolution implementation to ease the transition and reduce time to market (Optional but desirable).
\end{itemize}
\section{Related Work}
Since the popularization of IEEE 802.11, several papers have proposed modifications to the contention protocol that is used for sharing the medium.
They can be categorized in three groups regarding the approach they use.
The first one prevents that the contention window is reset to its minimum value after successful transmissions.
Examples of this first group include \cite{bharghavan1994map,wang2004ncr}.
This solution improves throughput in saturation conditions at the price of lowering short term fairness.

The second groups involves the accurate estimation of the number of contenders to adjust the contention parameters.
Two examples of this group are \cite{cali2000dti,lopez-toledo2006aoi}.
This approach offers some throughput and fairness gains at the expense of increased implementation complexity.
As the number of contenders is estimated relying on the number of collisions, the presence of channel errors further complicates the estimation.
Furthermore, there is a fundamental trade-off between the accuracy and reaction time of the estimation.

The aforementioned solutions are not able to fairly share the medium with legacy devices.In fact, these proposals are, generally speaking, less aggressive than the currently implemented protocol.
Consequently, in a shared network, the new stations would receive a smaller share of the available bandwidth in a mixed scenario.

A more important limitation of the solutions exposed so far is that the throughput is bounded by that of CSMA/CA with optimal configuration \cite{cali2000dti,bianchi2000pai}.
In this paper we focus on a third group of solutions that delivers throughput above the maximum attainable by CSMA/CA.

This third group of solutions uses a deterministic backoff after successful transmissions to further reduce the chances of collisions.
It was introduced in \cite{barcelo2008lba} and a more detailed analysis that includes both saturated and non-saturated conditions is presented in \cite{bellalta2009vtc}.
analyzed in depth in \cite{he2009srb}.

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{IEEEabrv,my_bib}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}
%
%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
%
%\end{thebibliography}
